Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-09-15 bbbq_article.tex
Remove MHCnuggets, fix #101

\begin{table}[]
  \begin{tabular}{llll}
    Goal & Tool & Reference \\ 
    \hline
    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Predict epitopes MHC-I and MHC-II & MHCnuggets                & \cite{shao2020high} \\
    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
    Call MHCnuggets from R            & mhcnuggetsr               & \cite{mhcnuggetsr} \\
    \verb;epitope-prediction; LUT     & epiprepreds               & \cite{epiprepreds} \\
    NetMHCIIpan LUT                   & nmhc2ppreds               & \cite{nmhc2ppreds} \\
    MHCnuggets LUT                    & mhcnpreds                 & \cite{mhcnpreds} \\
    Sequence alignment                & ClustalW                  & \cite{thompson1994clustal} \\
    Conservation scores               & msa                       & \cite{bodenhofer2015msa} \\
    Combine all                       & bbbq                      & \cite{bbbq}
  \end{tabular}
  \caption{
    Overview of all software used in this research.
    'LUT' is an abbreviation for 'Look-up table'. 
  }
  \label{table:software_used}
\end{table}

@@

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
A tool developed later is \verb;epitope-prediction; (\cite{bianchi2017}),
which uses a stabilized matrix method (\cite{kim2009derivation}).
The most recent FOSS tool available is 
MHCnuggets (\cite{shao2020high}) (which can do both MHC-I 
and MHC-II prediction) which is used in this study

@@

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
gave rise to the most accurate predictions.
Later tools are 
\richel{name of tool here} (\cite{zhang2013predicting}),
\richel{name of tool here} (\cite{trolle2014nettepi}),
\richel{name of tool here} (\cite{zhang2015accurate})
and the very recent MHCnuggets (\cite{shao2020high}),
the latter which is used in this study.

@@

The R programming language is used for the complete 
experiment, including the analysis.
The complete experiment is bundled in the 'bbbq' R package,
which is dependent on 'pureseqtmr', 
'mhcnuggetsr' and 'mhcnpreds',
'epitope-prediction' and 'epiprepreds',
'netmhc2pan' and 'nmhc2ppreds'
as described below.

@@

The R package 'mhcnuggetsr' is written to call the 
'MHCnuggets' Python package.
'mhcnuggetsr' allows to install 'MHCnuggets' and call it
to predict IC50 values for multiple MHC haplotypes that 'MHCnuggets' 
has already been trained upon.
Excluded is the feature to train 'MHCnuggets' on new data and/or other
haplotypes, as this study only needs IC50 predictions.
'mhcnuggetsr' has been submitted to CRAN.

To be able to use pre-calculated 
\verb;epitope-prediction;, NetMHCIIpan and MHCnuggets
predictions (i.e. a distribution of IC50 values per haplotype, 
derived from random peptides), the 'epiprepreds', 'nmhc2ppreds' 
and 'mhcnpreds' R package have been 
developed. In practice, these are nothing more than a big look-up table.
If \verb;epitope-prediction; is put onto CRAN, 'epiprepreds'
will be submitted as well.
If 'netmhc2pan' is allowed onto CRAN, 'nmhc2ppreds'
will be submitted as well.
If 'mhcnuggetsr' is put onto CRAN, 'mhcnpreds'
will be submitted as well.

@@

The earlier study uses a hand-crafted method, 
that has been trained on MHC-I haplotypes only.
This study uses MHCnuggets, which (1) free and open source, and (2) predicts
IC50s on both MHC-I and MHC-II.

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc1_grid.png}
  \caption{
    Percentage of MHC-I epitopes overlapping with TMHs
    for a human, viral and bacterial proteome.
    Epitope prediction by MHCnuggets.
  }
  \label{fig:1_mhcnuggets}
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc2_grid.png}
  \caption{
    \% epitopes overlapping with transmembrane helix
    for a human, SARS-CoV-2 (labeled 'covid') 
    and Mycobacterium tuberculosis (labeled 'myco') proteome,
    Epitope prediction by MHCnuggets.
  }
  \label{fig:2_mhcnuggets}
\end{figure}

@@

Per MHC haplotype, we simulated random peptides and measured the
predicted IC50 using MHCnuggets. 
Here, we show the IC50 value per haplotype that
is used to determine if a peptide binds to the haplotype's MHC.

>>>>

\begin{table}[]
  \begin{tabular}{llll}
    Goal & Tool & Reference \\ 
    \hline
    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
    \verb;epitope-prediction; LUT     & epiprepreds               & \cite{epiprepreds} \\
    NetMHCIIpan LUT                   & nmhc2ppreds               & \cite{nmhc2ppreds} \\
    Sequence alignment                & ClustalW                  & \cite{thompson1994clustal} \\
    Conservation scores               & msa                       & \cite{bodenhofer2015msa} \\
    Combine all                       & bbbq                      & \cite{bbbq}
  \end{tabular}
  \caption{
    Overview of all software used in this research.
    'LUT' is an abbreviation for 'Look-up table'. 
  }
  \label{table:software_used}
\end{table}

@@

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
A tool developed later is \verb;epitope-prediction; (\cite{bianchi2017}),
which uses a stabilized matrix method (\cite{kim2009derivation}).
The most recent and promising FOSS tool available now appears
to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
and MHC-II prediction.

@@

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
gave rise to the most accurate predictions.
Later tools are NetTepi (\cite{trolle2014nettepi}),
a genetic-algorithm-based ensemble method (\cite{zhang2015accurate})
and (again) the very recent MHCnuggets (\cite{shao2020high}).

@@

The R programming language is used for the complete 
experiment, including the analysis.
The complete experiment is bundled in the 'bbbq' R package,
which is dependent on 'pureseqtmr', 
'epitope-prediction' and 'epiprepreds',
'netmhc2pan' and 'nmhc2ppreds'
as described below.

@@

To be able to use pre-calculated 
\verb;epitope-prediction; and NetMHCIIpan
predictions (i.e. a distribution of IC50 values per haplotype, 
derived from random peptides), the 'epiprepreds' and 'nmhc2ppreds' 
R packages have been 
developed. In practice, these are nothing more than a big look-up table.
If \verb;epitope-prediction; is put onto CRAN, 'epiprepreds'
will be submitted as well.
If 'netmhc2pan' is allowed onto CRAN, 'nmhc2ppreds'
will be submitted as well.

@@

The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
that has been trained on MHC-I haplotypes only,
which is used here again. For MHC-II IC50 predictions, the
'netmhc2pan' R package is used, which calls NetMHC2pan.
Both are packages are free and open-source.

@@

@@

@@

Per MHC haplotype, we simulated random peptides and measured the
predicted IC50 using \verb;epitope-prediction; (for MHC-I) 
and netmhc2pan (for MHC-II). 
Here, we show the IC50 value per haplotype that
is used to determine if a peptide binds to the haplotype's MHC.
